---
title: "00_learn_ggmap"
author: "Author: [Steve, Yu](https://github.com/littlefish0331)"
date: "`r Sys.setlocale('LC_TIME', 'English'); format(Sys.time(), '%Y %b %d %a, %H:%M:%S')`" 
output:
  rmdformats::readthedown:
    css: style.css
    self_contained: TRUE
    thumbnails: FALSE
    lightbox: TRUE
    gallery: FALSE
    highlight: tango #探戈橘
    code_folding: show
    toc_depth: 3
    
---

```{r setup, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
rm(list = ls()); gc()
library(knitr)
library(kableExtra)
library(dplyr)
library(data.table)
library(ggmap)
knitr::opts_chunk$set(
	# 這邊是針對所有chunk的設定
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# reference

[dkahle/ggmap: A package for plotting maps in R with ggplot2](https://github.com/dkahle/ggmap)

---

# get location latlon range

[r - Is ggmap broken? Basic qmap() produces "arguments imply differing number of rows: 0,1" - Stack Overflow](https://stackoverflow.com/questions/52704695/is-ggmap-broken-basic-qmap-produces-arguments-imply-differing-number-of-rows)

```{r}
# library(tmaptools)
# geocode_OSM("Taiwan")
# $query
# [1] "Taiwan"
# 
# $coords
#         x         y 
# 120.98202  23.97394 
# 
# $bbox
#      xmin      ymin      xmax      ymax 
# 114.35932  10.37301 122.29700  26.43722 

# geocode_OSM("Hsinchu")
# $query
# [1] "Hsinchu"
# 
# $coords
#         x         y 
# 120.96868  24.80663 
# 
# $bbox
#      xmin      ymin      xmax      ymax 
# 120.87942  24.71257 121.03354  24.85321
```

---

# README

一些關於調用 google map 的注意事項。

This can be done on a temporary basis with register_google(key = "[your key]") or   
permanently using register_google(key = "[your key]", write = TRUE)  

(note: this will overwrite your ~/.Renviron file by replacing/adding the relevant line).  
If you use the former, know that you’ll need to re-do it every time you reset R.

```{r}
maps_static_api_key <- 
  read.table("E:/API_key/google_api_key.txt", header = T, stringsAsFactors = F) %>%
  setDT() %>% .[API_name=="maps_static", API_key]
register_google(key = maps_static_api_key)
# google_key()
# has_google_key()
```


## 簡單示例

stamenmap 是不用錢的

```{r}
# American
us <- c(left = -125, bottom = 25.75, right = -67, top = 49)
get_stamenmap(us, zoom = 4, maptype = "toner-lite") %>% ggmap()
get_stamenmap(us, zoom = 4, maptype = "terrain") %>% ggmap()
```

```{r}
# taiwan
tw <- c(left = 117.35932, bottom = 20.37301, right = 123.29700, top = 26.43722)
get_stamenmap(tw, zoom = 7, maptype = "toner-lite") %>% ggmap() 
get_stamenmap(tw, zoom = 7, maptype = "terrain") %>% ggmap() 
get_stamenmap(tw, zoom = 7, maptype = "terrain") %>% ggmap(extent = "device") #坐標軸會消失。
```

## qmplot

```{r}
library(forcats)

# define helper
`%notin%` <- function(lhs, rhs) !(lhs %in% rhs)

# reduce crime to violent crimes in downtown houston
violent_crimes <- crime %>% 
  filter(
    offense %notin% c("auto theft", "theft", "burglary"),
    -95.39681 <= lon & lon <= -95.34188,
     29.73631 <= lat & lat <=  29.78400
  ) %>% 
  mutate(
    offense = fct_drop(offense),
    offense = fct_relevel(offense, c("robbery", "aggravated assault", "rape", "murder"))
  )

# use qmplot to make a scatterplot on a map

qmplot(x = lon, y = lat, data = violent_crimes, 
       maptype = "toner-lite",
       color = I("red"))
# Using zoom = 14...
# Map tiles by Stamen Design, under CC BY 3.0. Data by OpenStreetMap, under ODbL.
```

## qmplot + ggplot2:density2d

All the ggplot2 geom’s are available.  
For example, you can make a contour plot with geom = "density2d"

```{r}
qmplot(x = lon, y = lat, data = violent_crimes[1:100, ], 
       maptype = "toner-lite",
       color = I("red"))

# 是點位的密集程度
qmplot(x = lon, y = lat, data = violent_crimes[1:100, ], 
       maptype = "toner-lite", geom = "density2d", 
       color = I("red"))
```

## qmplot + ggplot2:stat_density_2d, scale_fill_gradient2

In fact, since ggmap’s built on top of ggplot2,   
all your usual ggplot2 stuff (geoms, polishing, etc.) will work,   
and there are some unique graphing perks ggmap brings to the table, too.

```{r}
violent_crimes$offense %>% unique()
robberies <- violent_crimes %>% filter(offense == "robbery")
```

```{r}
qmplot(x = lon, y = lat, data = robberies, 
       maptype = "toner-lite",
       color = I("red"))
qmplot(x = lon, y = lat, data = robberies, 
       maptype = "toner-background",
       geom = "blank", #點位空白
       zoom = 14, #縮放
       darken = .7, #背景變暗
       legend = "topleft") +
  stat_density_2d(mapping = aes(fill = ..level..), geom = "polygon", alpha = .3, color = NA) +
  scale_fill_gradient2("Robbery\nPropensity", 
                       low = "white", mid = "yellow", high = "red", 
                       midpoint = 650)
#  Map tiles by Stamen Design, under CC BY 3.0. Data by OpenStreetMap, under ODbL.
```

qmplot + ggplot2:Facet

Faceting works, too:

```{r}
qmplot(x = lon, y = lat, data = violent_crimes, 
       maptype = "toner-background", 
       color = offense) + 
  facet_wrap(~ offense)
#  Using zoom = 14...
#  Map tiles by Stamen Design, under CC BY 3.0. Data by OpenStreetMap, under ODbL.
```

---

## Google Maps and Credentials

Source: 
- https://maps.googleapis.com/maps/api/staticmap
  - ?center=waco%20texas
  - &zoom=12
  - &size=640x640
  - &scale=2
  - &maptype=terrain
  - &key=xxx

- https://maps.googleapis.com/maps/api/geocode/json
  - ?address=waco+texas
  - &key=xxx

```{r}
get_googlemap("waco texas", zoom = 12, archiving = T) %>% ggmap()
get_googlemap("taiwan", zoom = 7, scale = 2, 
              maptype = "roadmap", archiving = T) %>% ggmap()
```

## Google map style

```{r}
get_googlemap("waco texas", zoom = 12, maptype = "terrain", archiving = T) %>% ggmap()
get_googlemap("waco texas", zoom = 12, maptype = "satellite", archiving = T) %>% ggmap()
get_googlemap("waco texas", zoom = 12, maptype = "hybrid", archiving = T) %>% ggmap()
get_googlemap("waco texas", zoom = 12, maptype = "roadmap", archiving = T) %>% ggmap()
```

Google's geocoding and reverse geocoding API's are available through geocode() and revgeocode(), respectively:

- geocode():
  - https://maps.googleapis.com/maps/api/geocode/json
  - ?address=1301+S+University+Parks+Dr,+Waco,+TX+76798
  - &key=xxx

```{r}
geocode("1301 S University Parks Dr, Waco, TX 76798")
## A tibble: 1 x 2
#      lon   lat
#    <dbl> <dbl>
#  1 -97.1  31.6
```

- revgeocode():
  - https://maps.googleapis.com/maps/api/geocode/json
  - ?latlng=31.55098,-97.1161
  - &key=xxx

```{r}
revgeocode(c(lon = -97.1161, lat = 31.55098))
#Multiple addresses found, the first will be returned:
#  55 Baylor Ave, Waco, TX 76706, USA
#  1437 FM434, Waco, TX 76706, USA
#  Bear Trail, Waco, TX 76706, USA
#  Robinson, TX 76706, USA
#  Waco, TX, USA
#  McLennan County, TX, USA
#  Texas, USA
#  United States
#[1] "55 Baylor Ave, Waco, TX 76706, USA"
```

There is also a mutate_geocode() that works similarly to dplyr’s mutate() function:

https://maps.googleapis.com/maps/api/geocode/json?address=white+house&key=xxx  
https://maps.googleapis.com/maps/api/geocode/json?address=waco+texas&key=xxx

```{r}
tibble(address = c("white house", "", "waco texas"))
tibble(address = c("white house", "", "waco texas")) %>% 
  mutate_geocode(address)
# "white house" not uniquely geocoded, using "1600 pennsylvania ave nw, washington, dc 20500, usa"

## A tibble: 3 x 3
#    address       lon   lat
#    <chr>       <dbl> <dbl>
#  1 white house -77.0  38.9
#  2 ""           NA    NA  
#  3 waco texas  -97.1  31.5
```

Treks use Google’s routing API to give you routes  
(route() and trek() give slightly different results; the latter hugs roads):

已經改為 directions api，要使用時再開起來。

- trek_df():  
  - https://maps.googleapis.com/maps/api/directions/json
  - ?origin=houson,+texas
  - &destination=waco,+texas
  - &key=xxx
  - &mode=driving
  - &alternatives=false
  - &units=metric

- qmap():
  - https://maps.googleapis.com/maps/api/staticmap
  - ?center=college%20station,%20texas
  - &zoom=8&size=640x640
  - &scale=2
  - &maptype=terrain
  - &language=en-EN
  - &key=xxx

  - https://maps.googleapis.com/maps/api/geocode/json
  - ?address=college+station,+texas
  - &key=xxx

```{r}
trek_df <- trek(from = "houson, texas", to = "waco, texas", structure = "route")
# route_df <- route(from = "houson, texas", to = "waco, texas", structure = "route")

qmap("college station, texas", zoom = 8) +
  geom_path(aes(x = lon, y = lat), 
            data = trek_df, 
            colour = "blue",
            size = 1.5, 
            alpha = .5,
            lineend = "round")
```

(They also provide information on how long it takes to get from point A to point B.)  
Map distances, in both length and anticipated time(預期時間), can be computed with mapdist().  
Moreover the function is vectorized:

distance matrix api，要使用時再開起來。

- mapdist(): 
  - https://maps.googleapis.com/maps/api/distancematrix/json
  - ?origins=dallas
  - &destinations=waco,+texas
  - &key=xxx
  - &mode=driving
  
  - https://maps.googleapis.com/maps/api/distancematrix/json
  - ?origins=houston,+texas
  - &destinations=waco,+texas
  - &key=xxx&mode=driving

```{r}
# mapdist(from = c("houston, texas", "dallas"), 
mapdist(from = c("houston, texas", "dallas"), 
        to = "waco, texas")
# # A tibble: 2 x 9
#    from          to               m    km miles seconds minutes hours mode  
#    <chr>         <chr>        <int> <dbl> <dbl>   <int>   <dbl> <dbl> <chr> 
#  1 houston, tex… waco, texas 298227  298. 185.    10257   171.   2.85 driving
#  2 dallas        waco, texas 152480  152. 94.8    5356    89.3   1.49 driving
```

---

# END

