---
title: "05_learn_heatmap_with_tmap"
author: "Author: [Steve, Yu](https://github.com/littlefish0331)"
date: "`r Sys.setlocale('LC_TIME', 'English'); format(Sys.time(), '%Y %b %d %a, %H:%M:%S')`" 
output:
  rmdformats::readthedown:
    css: style.css
    self_contained: TRUE
    thumbnails: FALSE
    lightbox: TRUE
    gallery: FALSE
    highlight: tango #探戈橘
    code_folding: show
    toc_depth: 3
    
---

```{r setup, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
rm(list = ls()); gc()
library(knitr)
library(kableExtra)
library(dplyr)
library(data.table)
library(tmap)
knitr::opts_chunk$set(
	# 這邊是針對所有chunk的設定
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# File Target

- learn tmp package

---

# tmp tutorial

有很多個，以官方提供的資料或連結為主，  
這是第一份。

[tmap: get started!](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html)

---

# tmap: get started!

With the tmap package, thematic maps can be generated with great flexibility.  
The syntax for creating plots is similar to that of ggplot2, but tailored to maps.  
This vignette is for those who want to get started with tmap within a couple of minutes. 
使用tmap包，可以非常靈活地生成主題地圖。  
創建圖的語法與ggplot2相似，但是是為圖量身定制的。  
此vignette適用於希望在幾分鐘之內開始使用tmap的人。

A more detailed description of tmap can be found in an [article](https://www.jstatsoft.org/article/view/v084i06) published in the Journal of Statistical Software ([JSS](https://www.jstatsoft.org/index)), which describes tmap version 1.11-2.  
The changes in version 2.0 are described in [`vignette("tmap-changes")`](https://cran.r-project.org/web/packages/tmap/doc/tmap-changes.html).
有關tmap的更詳細描述，可以在《統計軟件雜誌》(JSS)上發表的一篇文章中找到， 
該文章介紹了tmap版本1.11-2。  
vignette("tmap-changes")中介紹了2.0版中的更改。

---

## Hello World!

A good place to start is to create a map of the world.  
After [installing](https://github.com/mtennekes/tmap#installation) tmap,  
the following lines of code should create the map shown below:

```{r}
data("World")

tm_shape(shp = World) +
    tm_polygons(col = "HPI")
```

The object `World` is a spatial object of class `sf` from the [sf package](https://cran.r-project.org/package=sf);  
it is a `data.frame` with a special column that contains a geometry for each row, in this case polygons.  

對象`World`是[sf包](https://cran.r-project.org/package=sf)中類`sf`的空間對象。  
它是一個帶有特殊列的`data.frame`，其中每一行包含幾何圖形，在本例中為多邊形。 

In order to plot it in tmap, you first need to specify it with `tm_shape`.  
Plotting layers can be added with the `+` operator, in this case `tm_polygons`.  
There are many layer functions in tmap, which can easily be found in the documentation  
by their `tm_` prefix. See also `?'tmap-element'`.  
為了在tmap中繪製它，首先需要用`tm_shape`指定它。 可以使用+運算符(在本例中為tm_polygons)添加繪圖層。  tmap中有很多圖層函數，可以在文檔中通過它們的`tm_`前綴輕鬆找到。  
另見`?'tmap-element'`。

> 可以去看一下 worldy資料集 的 geometry 欄位。


## Interactive maps

Each map can be plotted as a static image or viewed interactively  
using `"plot"` and `"view"` modes, respectively.  
每個地圖可以繪製為靜態圖像，也可以分別使用"繪圖"和"查看"模式進行交互查看。  

The mode can be set with the function `tmap_mode`,  
and toggling between the modes can be done with the ‘switch’ `ttm()`.  
可以使用功能`tmap_mode`設置模式，並可以通過 switch `ttm()`來切換模式。

> ttm(): 不用參數，執行過後，就表示tmap_mode()從 plot 轉換成 view。

```{r}
tmap_mode(mode = "view")

tm_shape(shp = World) +
    tm_polygons(col = "HPI")
```


## Multiple shapes and layers

A shape is a spatial object (with a class from `sf`, `sp` `stars`, or `raster`).  
Multiple shapes and also multiple layers per shape can be plotted:

形狀是空間對象(具有`sf`, `sp` `stars`, or `raster`的類)。  
可以繪製多個形狀，每個形狀也可以繪製多個圖層: 

```{r}
# elevation: 海拔
# metro: 地鐵。資料集為各城市的地鐵運量
# scale: symbol size multiplier number.
# is.master = T，地圖會預先縮到這個圖層。
# 似乎沒辦法讓 legend 個別設置是否要顯示。

# ---
# 在 plot mode 之下，tm_legend(show = FALSE)，會將所有legend隱藏。
# 在 view mode 之下，tm_legend(show = TRUE)，不會顯示圖形(以此例為symbols)圖例，但是區塊圖例會顯示。

# ---
# data(World, metro, rivers, land)
data(World, metro, land)

tmap_mode("plot")
## tmap mode set to plotting

# ttm()+
tm_shape(shp = land) +
  tm_raster(col = "elevation", palette = terrain.colors(10), legend.show = T) +
  
  tm_shape(shp = World) +
  tm_borders(col = "white", lwd = .5) +
  tm_text(text = "iso_a3", size = "AREA") + 
  
  tm_shape(shp = metro) + 
  tm_symbols(col = "red", size = "pop2020", scale = .5) +
  tm_legend(show = F)
```


## Facets

Facets can be created in three ways:
可以通過以下三種方式創建構面:

1. By assigning multiple variable names to one aesthetic:  
通過為一種美學分配多個變量名稱：

> 關於同步的 red circle，會固定在經緯度(0,0)的地方，有點礙眼但是也沒有辦法， 問這是JS的問題。
> 不過把地圖拉大的話，就看不見了~

[Red circles in latticeView sync mode · Issue #51 · r-spatial/mapview](https://github.com/r-spatial/mapview/issues/51)

> 另外就是 red circle 無法在圖層的上面，目前還沒有找到辦法。

```{r}
tmap_mode("view")
tm_shape(shp = World) +
  tm_polygons(col = c("HPI", "economy")) +
  tm_facets(sync = TRUE, ncol = 2)
```

```{r eval = FALSE}
# 其他地圖同步的寫法
library(leaflet)
library(mapview)
l1 <- leaflet() %>% addProviderTiles(provider = "CartoDB.Positron")
l2 <- leaflet() %>% addProviderTiles(provider = "OpenStreetMap")
sync(l1, l2)

m1 <- mapview()
m2 <- mapview()
sync(m1, m2, sync.cursor = T)
```

2. By splitting the spatial data with the `by` argument of `tm_facets`.  
通過使用`tm_facets`的`by`參數拆分空間數據。

```{r}
# 如果 facets 的數量沒有正確可能是套件一開始有限制圖塊數量
# 避免電腦跑不出來~
# default: c(facets.view = 4, facets.plot = 64)
# tmap_options(limits = c(facets.view = 12))
# tmap_options(limits = c(facets.plot = 12))

# ---
tmap_mode("plot")
## tmap mode set to plotting

data(NLD_muni)
NLD_muni$perc_men <- NLD_muni$pop_men / NLD_muni$population * 100

tm_shape(NLD_muni) +
    tm_polygons(col = "perc_men", palette = "RdYlBu") +
    tm_facets(by = "province")
```

3. By using the `tmap_arrange` function:  
通過使用`tmap_arrange`函數:

> legend 大小的相關設置，除了在圖層函數(tm_polygons, tm_bubbles)，  
> 本身就有之外，tm_layout()也有更多的設定。

```{r}
tmap_mode("plot")
## tmap mode set to plotting

data(NLD_muni)
tm1 <- tm_shape(NLD_muni) + tm_polygons(col = "population", convert2density = TRUE)
tm2 <- tm_shape(NLD_muni) + tm_bubbles(size = "population") + tm_layout(legend.show = T)

tmap_arrange(tm1, tm2)
## Legend labels were too wide.  
## Therefore, legend.text.size has been set to 0.27.  
## Increase legend.width (argument of tm_layout) to make the legend wider 
## and therefore the labels larger.

## The legend is too narrow to place all symbol sizes.
```


## Basemaps and overlay tile maps

底圖和疊加圖塊地圖

Tiled basemaps can be added with the layer function `tm_basemap`.  
Semi-transparent overlay maps (for example annotation labels) can be added with `tm_tiles`.

可以使用圖層功能`tm_basemap`添加平舖的底圖。  
可以使用`tm_tiles`添加半透明的覆蓋圖(例如註釋標籤)。

```{r}
# alpha: 透明度。

# ---
tmap_mode(mode = "view")

# tm_basemap(server = NA) +
# tm_basemap(server = NULL) +
tm_basemap(server = "Stamen.Watercolor") +
  tm_shape(shp = metro) + 
  tm_bubbles(size = "pop2020", col = "red") +
  tm_tiles(server = "Stamen.TonerLabels", alpha = 0.5)
```

See a [preview of the available tilemaps](http://leaflet-extras.github.io/leaflet-providers/preview/).  
This list is also accessible in R: `leaflet::providers`.


## Options and styles

The functions `tm_layout` and `tm_view` are used to specify the map layout and the interactive aspects respectively.  
函數`tm_layout`和`tm_view`分別用於指定地圖佈局和交互方面。  

These functions can be used in the same way as the layer functions, e.g.  
這些功能可以與圖層功能相同的方式使用，例如

```{r}
# inner.margins: 地圖和邊框的距離，是指地圖和黑色的邊框。bottom, left, top, and right。

# ---
tmap_mode("plot")
## tmap mode set to plotting

tm_shape(World) +
  tm_polygons("HPI") +
  tm_layout(bg.color = "skyblue", 
            inner.margins = c(0, .02, .02, .02))
```

These options, as well as a couple of others, can also be set within with `tmap_options`,  
which works in the same way as the base R function `options`.  
這些選項以及其他幾個選項也可以在`tmap_options`中設置，  
其作用方式與基本R函數`options`相同。  

The main advantage is that these options are set globally,  
so they do not have to be specified in each map, for the duration of the session.
主要優點是這些選項是全局設置的，  
因此在會話期間不必在每個映射中都指定它們。

```{r}
tmap_options(bg.color = "black", legend.text.color = "white")

tm_shape(World) +
  tm_polygons(col = "HPI", 
              title = "Happy Planet Index")
```

A style is a certain configuration of the tmap options.  
樣式是tmap選項的特定配置。

```{r}
tmap_style("classic")
## tmap style set to "classic"
## other available styles are: 
## "white", "gray", "natural", "cobalt", "col_blind", 
## "albatross", "beaver", "bw", "watercolor"

tm_shape(World) +
    tm_polygons("HPI", title = "Happy Planet Index")
```

可以透過 tmap_options_diff() 觀察 options() 有哪些地方變動。

```{r}
# see what options have been changed
# diff的部分是來自 tmap_style("classic") 這個指令。
tmap_options_diff()
## current tmap options (style "classic") that are different from default tmap options (style "white"):
## $sepia.intensity
## [1] 0.7
## 
## $frame.double.line
## [1] TRUE
## 
## $fontfamily
## [1] "serif"
## 
## $compass.type
## [1] "rose"
## 
## $basemaps
## [1] "Esri.WorldTopoMap"
## 
## $basemaps.alpha
## [1] 0.5
```

```{r}
# reset the options to the default values
tmap_options_reset()
## tmap options successfully reset
```

New styles can be created; see `?tmap_options`.


## Exporting maps

```{r}
tm <- tm_shape(World) +
  tm_polygons("HPI", legend.title = "Happy Planet Index")

## save an image ("plot" mode)
tmap_save(tm = tm, filename = "../data/export_map/world_map.png")

## save as stand-alone HTML file ("view" mode)
## 這個一定要用絕對路徑，套件沒有寫好。
## 如果只是存在當前的目錄底下(不可以再多資料夾)則可使用為相對路徑。
## 如果不在當前目錄，會多一個 world_map_files/
# tmap_save(tm, filename = "E:/NCHC/project2020/heatmap_learn/data/export_map/world_map.html")
tmap_save(tm, filename = "world_map.html")
```


## leaflet and shiny integration

Maps made with tmap can be exported to leaflet with the function `tmap_leaflet`.  
使用tmap製作的地圖可以使用tmap_leaflet函數導出到leaflet。  

Hence, it is possible to make use of leaflet own functionalities.  
This function is also needed when using tmap in shiny:
因此，可以利用leaflet自身的功能。  
在shiny中使用tmap時，也需要此功能：


```{r eval = FALSE}
# in UI part:
leafletOutput("my_tmap")

# in server part
output$my_tmap = renderLeaflet({
    tm <- tm_shape(World) + 
      tm_polygons("HPI", legend.title = "Happy Planet Index")
    tmap_leaflet(tm)
})
```

## Quick thematic map

Maps can also be made with one function call. This function is `qtm`:

```{r}
# 會依循上一次的 tmap_mode()
qtm(shp = World, fill = "HPI", fill.pallete = "RdYlGn")
```

## Tips 'n Tricks

Run:  
有很多個小販力可以估看，  
重複執行 tmap_tip() 就可以看到程式碼跟效果了

```{r}
tmap_tip()
```

---

# END

