---
title: "06_learn_heatmap_with_tmap"
author: "Author: [Steve, Yu](https://github.com/littlefish0331)"
date: "`r Sys.setlocale('LC_TIME', 'English'); format(Sys.time(), '%Y %b %d %a, %H:%M:%S')`" 
output:
  rmdformats::readthedown:
    css: style.css
    self_contained: TRUE
    thumbnails: FALSE
    lightbox: TRUE
    gallery: FALSE
    highlight: tango #探戈橘
    code_folding: show
    toc_depth: 3
    
---

```{r setup, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
rm(list = ls()); gc()
library(knitr)
library(kableExtra)
library(dplyr)
library(data.table)
library(tmap)
knitr::opts_chunk$set(
	# 這邊是針對所有chunk的設定
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# File Target

- learn tmp package

---

# tmp tutorial

有很多個，以官方提供的資料或連結為主，  
這是第二份。  

[tmap: version changes](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-changes.html)

---

# tmap: version changes

This vignette summarizes the main changes in tmap 2.x/3.x in comparison to tmap 1.11-2,  
which is described in detail in the [JSS paper](https://www.jstatsoft.org/article/view/v084i06).  
該vignette總結了與tmap 1.11-2相比，tmap 2.x/3.x的主要變化， 
JSS論文對此進行了詳細介紹。  

The code of this paper for version 2.x/3.x can be found in [`vignette("tmap-JSS-code")`](https://cran.r-project.org/web/packages/tmap/doc/tmap-JSS-code.html).
可以在vignette("tmap-JSS-code")中找到針對2.x/3.x版的本文代碼。  

For people who are new to tmap, see [`vignette("tmap-getstarted")`](https://cran.r-project.org/web/packages/tmap/doc/tmap-getstarted.html).  
對於不熟悉tmap的人，請參見vignette("tmap-getstarted")。

---

## Function renaming: tm_ and tmap_

As of version 2.0, all tmap functions have the prefix `tm_` or `tmap_`,  
with the exception of the function `qtm`.  
從2.0版開始，所有tmap函數均具有前綴tm_或tmap_，  
但函數qtm除外。  

Just like in tmap version 1.x, the layer functions start with `tm_`.  
就像在tmap版本1.x中一樣，圖層功能以tm_開頭。 

In tmap 1.x, the names of the other functions where inconsistent,  
e.g. animation_tmap and tmap_arrange. In tmap 2.x/3.x, these functions are prefixed with tmap_.  
在tmap 1.x中，其他函數的名稱不一致，  
例如 animation_tmap和tmap_arrange。在tmap 2.x/3.x中，這些函數以tmap_為前綴。  

For instance animation_tmap has been renamed to tmap_animation.
例如，animation_tmap已重命名為tmap_animation。


## Spatial data classes

空間數據類

tmap 1.x uses `sp` for representing vector data, and `raster` for raster data.  
tmap 2.x uses `sf` for representing vector data, and `raster` for raster data.  
tmap 3.x uses `sf` for representing vector data, and `stars` for raster data.

說明不同版本下，對於空間資料使用哪種套件做包裝。

`sf` (simple features) objects have a simpler structure than `sp` objects.  
An sf object is nothing more than a `data.frame` with a special geometry column  
that contains the geometries for the corresponding rows.  
sf(簡單特徵)對象的結構比sp對像簡單。  
sf對像不過是帶有特殊幾何列的data.frame，其中包含對應行的幾何。  

Such a geometry can be of type spatial point(s), line(s) or polygon(s)  
or any combination of these in a ‘geometrycollection’ (see `vignette("sf1")`).  
這樣的幾何可以是"geometrycollection"中的空間點，線或多邊形  
或它們的任意組合(請參見vignette(" sf1"))。

The layers functions, such as `tm_polygons`, will only draw what they are supposed to draw (in this case polygons).  
The newly added layer function `tm_sf` will draw all geometries.  
圖層功能(例如tm_polygons)將僅繪製應繪製的內容(在本例中為多邊形)。 新添加的圖層函數tm_sf將繪製所有幾何圖形。

Finally, all data objects in tmap (except `land`) have been transformed into `sf` objects.  
The object `land` has been transformed into a `stars` object.
最後，tmap中的所有數據對象(land除外)都已轉換為sf對象。  
land對已被轉換為stars對象。


## tm_tiles

Layer functions have been added to facilitate tile layers.  
In version 1.x, it was already possible to add basemaps.  
In version 2.0, this is facilitated by the explicit layer functions `tm_basemap` and `tm_tiles`.  
添加了圖層功能以方便平鋪圖層。  
在1.x版中，已經可以添加底圖。 
在版本2.0中，顯式圖層函數`tm_basemap`和`tm_tiles`促進了此操作。  

The former creates a basemap tile and the latter an overlay tile.  
The difference is that `tm_basemap` layers are always drawn at the bottom, whereas the plotting order of (overlay) `tm_tiles` layers is derived from the plot call.  
前者創建底圖圖塊，而後者創建覆蓋圖塊。  
區別在於，`tm_basemap`圖層始終在底部繪製，而(overlay)tm_tiles圖層的繪製順序是從繪圖調用中得出的。  

Overlay layers should be semi-transparent to be useful.  
覆蓋層應該是半透明的，這樣才有用。

Tip: all tilemaps from [http://leaflet-extras.github.io/leaflet-providers/preview/](http://leaflet-extras.github.io/leaflet-providers/preview/) can be used.  
The names of these maps can also be found in the list object `leaflet::providers`.

提示: 可以使用 http://leaflet-extras.github.io/leaflet-providers/preview/ 中的所有圖塊。  
這些映射的名稱也可以在列表對象`leaflet::providers`中找到。

```{r}
data(World, metro)
tmap_mode("view")

tm_basemap(server = leaflet::providers$CartoDB.PositronNoLabels, group = "CartoDB basemap") +
  
  tm_shape(shp = World) +
  tm_polygons(col = "HPI", group = "Countries") +
  
  tm_tiles(server = leaflet::providers$CartoDB.PositronOnlyLabels, group = "CartoDB labels") +
  
  tm_shape(metro) +
  tm_dots(col = "red", group = "Metropolitan areas")
```


## tmap options, styles, and formats

There are many options in tmap.  
In version 1.x, the default values were stored in `tm_layout` and `tm_view`.  
Wrapper functions like `tm_style_gray` and `tm_format_World` where used to change these options.
tmap中有很多選項。  
在版本1.x中，默認值存儲在`tm_layout`和`tm_view`中。  
包裝函數(如`tm_style_gray`和`tm_format_World`)用於更改這些選項。

In version 2.0, the options are stored in a similar fashion to the global options.  
A list of tmap options can be retrieved and options can be set with the function `tmap_options`.  
在2.0版中，選項以與全局選項相似的方式存儲。  
可以檢索tmap選項的列表，並可以使用功能`tmap_options`設置選項。  

Its behavior is similar to the base function `options`.  
其行為類似於基本功能選項。

A style is a configuration of the options used to style a map in a specific way.  
When set with `tmap_style`, the current options are overwritten according to the new style.  
The changes can be seen with `tmap_options_diff` and can be reset with `tmap_options_reset`.
樣式是用於以特定方式設置地圖樣式的選項的配置。  
使用`tmap_style`設置時，當前選項將根據新樣式覆蓋。  
更改可以通過`tmap_options_diff`查看，也可以通過`tmap_options_reset`重置。


```{r}
data(World)

# tmap_mode("plot") + 
qtm(World, fill = "life_exp")
```

```{r}
# 個人覺得不錯的: classic, gray, natural, cobalt, col_blind, watercolor

# ---
tmap_style(style = "classic")
## tmap style set to "classic"
## other available styles are: 
## "white", "gray", "natural", "cobalt", "col_blind", 
## "albatross", "beaver", "bw", "watercolor"

qtm(World, fill = "life_exp")
```

```{r}
tmap_options_diff()
## current tmap options (style "classic") that are different from default tmap options (style "white"):
## $sepia.intensity
## [1] 0.7
## 
## $frame.double.line
## [1] TRUE
## 
## $fontfamily
## [1] "serif"
## 
## $compass.type
## [1] "rose"
## 
## $basemaps
## [1] "Esri.WorldTopoMap"
## 
## $basemaps.alpha
## [1] 0.5

tmap_options_reset()
## tmap options successfully reset
```

Styles included in tmap are: "white", "gray", "natural", "cobalt", "col_blind", "albatross", "beaver", "bw", "classic".  
tmap中包含的樣式為："white", "gray", "natural", "cobalt(鈷)",
"col_blind", "albatross(信天翁)", "beaver(海狸)", "bw", "classic"。  

New styles can saved and loaded with `tmap_style_save` and `tmap_style_load` respectively.
可以分別使用`tmap_style_save`和`tmap_style_load`保存和加載新樣式。

In version 1.x, tmap contained a couple of predefined format functions, such as tm_format_World.  
In version 2.0, these have been replaced by the general function tm_format.  
With this function, the available formats can be retrieved, as well as the configuration of a specific format.  
With the function `tm_format_add` new formats can be created.  
在1.x版中，tmap包含幾個預定義的格式函數，例如tm_format_World。  
在2.0版中，這些已被通用函數`tm_format`代替。  
使用此功能，可以檢索可用格式以及特定格式的配置。  
使用功能`tm_format_add`可以創建新格式。

```{r}
tmap_format()
## available formats are: "World", "World_wide", "NLD", "NLD_wide"

panorama <- tmap_format(format = "World")
panorama$asp <- 6
tmap_format_add(panorama, name = "panorama")
## format panorama succesfully added. Use this format with tm_format("panorama")
## panorama: 全景圖

tmap_format()
## available formats are: "World", "World_wide", "NLD", "NLD_wide", "panorama"
```

```{r}
## NLD, National Levee Database: 國家堤防數據庫

# ---
tm_shape(World) + tm_polygons("HPI") + tm_format("panorama")
# tm_shape(World) + tm_polygons("HPI") + tm_format("NLD")
# tm_shape(World) + tm_polygons("HPI") + tm_format("NLD_wide")
# or: qtm(World, fill = "HPI", format = "panorama")
```


## color palettes

Viridis palettes (from the `viridis`/`viridisLite` package) are supported natively.  
E.g. `qtm(World, "HPI", fill.palette = "-plasma")`.  
本機支持Viridis調色板(來自viridis/viridisLite軟件包)。  
例如。 `qtm(World, "HPI", fill.palette = "-plasma")`。  

Notice that, like with the color brewer palette names, a minus sign will reverse the palette.  
See `tmaptools::palette_explorer()` to explore the color brewer and viridis palettes interactively.
請注意，與顏色調製器調色板名稱一樣，減號將反轉調色板。  
請參閱tmaptools::palette_explorer()以交互方式瀏覽顏色釀造商和viridis調色板。

> tmaptools::palette_explorer()，這東西方便好用，選好想要的顏色就會提供程式碼。

The arguments that define the color palette mapping has been made more intuitive. The newly added layer function arguments `midpoint` and `stretch.palette` replace `auto.palette.mapping`:  
定義調色板映射的參數已變得更加直觀。  
新添加的圖層函數自變量`midpoint`和`Stretch.palette`替換了`auto.palette.mapping`: 

- Categorical palettes.  
The argument `stretch.palette` determines whether a palette is stretched out when there are more categories than available colors.  
If set to `FALSE`, the palette is repeated.
分類調色板。  
參數Stretch.palette確定當類別多於可用顏色時是否擴展調色板。  
如果設置為FALSE，則重複調色板。

- Diverging palettes.  
The argument `midpoint` determines which value should be mapped to the middle (neutral) color of the diverging color palette.  
By default, it is set to 0 if there are positive and negative values.  
This could mean that only a part of the color palette is used,  
e.g. when the value range is \[-5, 20\] and the palette is `"RdBu"`, the colors will range from light red (-5) to white (0) to dark blue (20).  
If there are only positive or only negative values, the full palette is shown.  
不同的調色板。  
中點參數確定應將哪個值映射到發散調色板的中間(中性)顏色。  
默認情況下，如果有正值和負值，則將其設置為0。  
這可能意味著僅使用了調色板的一部分，  
例如，當值的範圍是[-5，20]並且調色板是"RdBu"時，顏色的範圍從淺紅色(-5)到白色(0)到深藍色(20)。  
如果只有正值或只有負值，則會顯示完整的調色板。

- Sequential palettes.  
Nothing has changed for sequential palettes.  
When `midpoint` has been specified, the palette is regarded as a diverging palette,  
with the middle color interpreted as the neutral color.  
順序調色板。  
順序調色板沒有任何變化。  
當指定了中點時，調色板被視為發散調色板，中間顏色被解釋為中性顏色。

> 顏色配置這一段，第二點大概懂，其他的可能要實際應用才知道。


## filter, difference between NA and NULL

> 欄位篩選，以及數據 NULL 和 NA 的差異

The argument filter has been added to `tm_shape`.  
Features that are included will be visualized as usual.  
參數過濾器已添加到tm_shape。  
所包含的功能將照常顯示。  

Excluded features will still be visualized but without color being mapped.  
Instead, they will be colored with `colorNULL`, a new argument of the layer functions.  
排除的特徵仍將可視化，但不會映射顏色。  
取而代之的是，它們將使用`colorNULL`著色，這是圖層函數的新參數。  

Note the difference with `colorNA`, which is used to color features with missing data.
注意colorNA的區別，`colorNA`用於為缺少數據的要素上色。


To show the potiential of this filter  
the following code chunk creates a map in which only European countries are colored:
為了顯示此過濾器的潛力，  
以下代碼塊創建了一個僅對歐洲國家進行著色的地圖：

```{r}
# id: name of the data variable that specifies the indices of the polygons. Only used for "view" mode 

# ---
# ttm()+
tm_shape(World, filter = World$continent=="Europe") +
    tm_polygons(col = "HPI", id = "name", 
                colorNULL = "red", #為排除的地區上色。
                colorNA = "green" #篩選出的地區，如果缺失值(NA)的顏色。
                )
```

Countries outside Europe are colored with `colorNULL`, which is `grey95` by default.  
It can be adjusted in the layer functions, such as `tm_polygons`, and in `tmap_options`.
歐洲以外的國家/地區以colorNULL上色，默認為grey95。  
可以在圖層函數(例如`tm_polygons`)和`tmap_options`中對其進行調整。

There are also missing values in the data (Kosovo and Moldova) which are colored with the `colorNA` (by default `grey85`).  
數據中還缺少值(科索沃和摩爾多瓦)，這些值用`colorNA`著色(默認為`grey85`)。


---

# END

